---
title: "Assignment 2: Base and Product Line Pricing Analysis"
author: "Günter J. Hitsch"
date: "April 14, 2020"
output:
  pdf_document:
    number_sections: yes
    toc: yes
urlcolor: blue
graphics: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, eval = TRUE,
                      fig.width = 4.5, fig.height = 3, fig.align = "right")
```

\setlength{\parskip}{6pt}
\newpage




# Overview

Your marketing-analytic consulting task is to conduct a base pricing analysis for P&G's flagship laundry detergent brand, Tide. The main questions to address are:

1. What is the extent of cannibalization within the Tide product line? 
2. Does Tide face a competitive threat from Wisk? 
3. How do you evaluate the current pricing tactics? Do you recommend changes?

You have access to scanner data in the laundry detergent category across 86 stores of a retail chain in Chicago. The data are in the file `Detergent.RData`. The data include weekly sales and price information for three products --- Tide 128 oz, Tide 64 oz, Wisk 64 oz --- across the 86 stores. The data are available for up to 300 weeks (the exact number of weeks included in the data varies across stores).

The variables in the data set are:

\medskip

-------------   -----------------------------------------------------
`store`         Store id number
`week`          Week
`acv`           ACV (all commodity volume), in $1,000
`promoflag`     = 1 if any product in the category was on promotion
`q_tide128`     Tide 128 oz: unit sales
`p_tide128`     Tide 128 oz: price ($)
`q_tide64`      Tide 64 oz: unit sales
`p_tide64`      Tide 64 oz: price ($)
`q_wisk64`      Wisk 64 oz: unit sales
`p_wisk64`      Wisk 64 oz: price ($)
-------------   -----------------------------------------------------

\medskip

Note that Wisk was discontinued in 2017 in favor of Persil.

\newpage




# Data description

Report the revenue market shares of the three products (percentage of total sales revenue across all store/week observations), and report the mean, median, and standard deviation of prices for the three products across store-weeks.

Then calculate the price gap (price difference) between (i) Tide 128oz and Tide 64oz and (ii) Tide 64oz and Wisk 64oz. Show histograms of the price gaps.

What do you learn from the summary statistics and price gap histograms about the ability to estimate demand models for the products?
```{r}
    library(ggplot2)
    library(tidyverse)
    library(dplyr)
    load("Detergent.RData")
```

For market share analysis
```{r}
    detergent_df %>%
    summarize(sales_tide128 = sum(q_tide128*p_tide128),
            sales_tide64 = sum(q_tide64*p_tide64),
            sales_wisk64 = sum(q_wisk64*p_wisk64),
            sales_total = sum(q_tide128*p_tide128)+sum(q_tide64*p_tide64)+sum(q_wisk64*p_wisk64),
            sales_tide128_pct = sales_tide128/sales_total,
            sales_tide64_pct = sales_tide64/sales_total,
            sales_wisk64_pct = sales_wisk64/sales_total)
    
    
    #tide 128 market share: 0.5685699
    #tide 64 market share: 0.2633647	
    #wisk 64 market share: 0.1680654
```

```{r, fig.width = 6, fig.height = 4, fig.align = "center"}
    Tide_diff_df = detergent_df
    Tide_diff_df$p_diff_tide128 = Tide_diff_df$p_tide128 - Tide_diff_df$p_tide64
    Tide_diff_df$p_diff_wisk64 = Tide_diff_df$p_wisk64 - Tide_diff_df$p_tide64
    
    Tide_diff_df = Tide_diff_df[,c("week", "p_diff_tide128",  "p_diff_wisk64")] %>% 
    pivot_longer(cols = p_diff_tide128:p_diff_wisk64,
                    names_to = c("Type of Value", "Type of Date", "Diff_from_Product"), names_sep = "_")
    
    Tide_diff_df = Tide_diff_df[,c("week", "Diff_from_Product",  "value")]
    
    p <- ggplot(Tide_diff_df, aes(x=value, color=Diff_from_Product)) + geom_histogram(fill="white")
    print(p)
```

```{r}   
    Tide_diff_df %>% 
      group_by(Diff_from_Product) %>% 
      summarize(mean_price_diff = mean(value), sd_price_diff = sd(value))
    # We can see that wisk is a little bit cheaper than tide in most cases.
    # tide128 is higher than 64 oz about 4 dollars for most cases.
```

It appears that the price of 128oz Tide varies in a broader range than the other two products. Our intuition suspects this could be due to promotions or geographic difference. 

It appears that the price gap between any two given products falls within a certain range, and we believe brand managers deliberately manage the price gap as it is part of their pricing strategy. The fluctuations could be a result of promotions and the different locations. Another interesting analysis to do would be review the price gap by store or chain, and how it evolves over time.
Response to question regarding our ability to estimate demand models:
There are 44235 observations (by product) in our statistics and we capture a range of different price points by product. We should be able to build a log-liner demand model with our data.

The standard deviation of the prices are within $1 for Tide 128oz and $0.5 for the other two products. Most stores keep a $4 price gap between Tide 128oz and Tide 64oz, and similar price for Tide 64oz and Wisk 64oz. Therefore, the data is good to estimate demand given the relatively stable price with established price gaps.

It is a good set for research.

\medskip

Hints:

(a) The revenue market shares can be easily calculated from the `detergent_df` table directly. Optionally, however, and as an exercise to hone your dplyr skills, you may convert the data to long format, where the price and quantity data are stacked on top of each other separately for each product. You can create such a long table as follows: 

    ```{r, eval = FALSE}
#detergent_df_long = 
detergent_df %>%
   pivot_longer(cols = q_tide128:p_wisk64,
                    names_to = c("variable", "product"), names_sep = "_") %>%
   pivot_wider(id_cols = c(product, store, week),
                   names_from = variable, values_from = value)
```

    This is beyond the scope of this assignment, but if you want to understand what the code above achieves, consult the *Reshaping tables (pivoting)* section in the R data science guide and the online documentation for pivoting in the `tidyr` package.
 
    Using the data in long format, `detergent_df_long`, you can calculate the *revenue market shares* more easily using `group_by` and `summarize`.
  
(b) Calculate the *price* summary statistics using `group_by` and `summarize`.

\newpage




# Demand estimation

Recall the definition of the *sales velocity*:
$$\text{velocity}=\frac{\text{unit sales}}{\text{ACV}}.$$

What is the purpose of dividing unit sales by ACV to construct the dependent variable?

The sales velocity tells how well a product really sells. The sales velocity looks at the sales rate with the ratio of the unit sales to the all commodity volumn. The dependent variable is separated from the inflation, or when the ACV fluctuates. The sales velocity is a better variable to compare between different products, retailers and sales channels within the same market.

## Basic demand model

Estimate log-linear demand models for the two Tide products (128 and 64 oz) by regressing the log of velocity on all prices (own and competing products). 

Discuss whether the demand estimates make sense. Are the magnitudes and signs of the estimated parameters as you would expect?

```{r, fig.width = 6, fig.height = 4, fig.align = "center"}

tide128lm = glm(log(q_tide128/acv) ~ log(p_tide128) + log(p_tide64) +log(p_wisk64), data = detergent_df)
summary(tide128lm)

tide64lm = glm(log(q_tide64/acv) ~  log(p_tide64) + log(p_tide128) + log(p_wisk64), data = detergent_df)
summary(tide64lm)

detergent_df %>% 
  select("store":"p_tide64") %>% 
  pivot_longer(cols = q_tide128:p_tide64, names_to = c("variable", "product"), names_sep = "_") %>% 
  pivot_wider(id_cols = c(product, store, week, acv),
                   names_from = variable, values_from = value)  %>%
  mutate(velocity = q/acv) %>% 
  ggplot(data = ., aes(x = log(p), y = log(velocity), color = product)) + 
  geom_point(alpha = 0.5, size = 0.1) + 
  geom_smooth(method = "lm", se = FALSE)
#It makes sense, cheaper sales faster

```

With 1% of the Tide 128 oz price increase and others fixed, the sales velocity for the Tide 128 oz would decrese by -4.597%. The price elasticity for Tide 128 oz is -4.597. The demand is elastic, as expected. But the value is higher or more elastic than expected. The customers can purchase the substitutes with similar unit price (price per oz), such as the Tide 64 oz or the Wisk 64 oz.

With 1% of the Tide 64 oz price increase and others fixed, the sales velocity for the Tide 128 oz would increase by 0.287%. With 1% of the Wisk 64 oz price increase and others fixed, the sales velocity for the Tide 128 oz would increase by 0.151%. The estimates are as expected. When the price for the smaller volumn products increase, the substitutes are either with same brand but different size or with same size but different brand. The sales for both kind of substitutes would increase (positive sign for the coefficient estimate for the Tide 128 oz sales velocity). But the customers might purchase the similar size over the same brand different size for similar purchase cycles or for the price decrease for the same product, which leads to lower estimate coefficients for the cross elasticity for the Tide 128 oz.

\medskip

Note: Instead of explicitly calculating sales velocity columns you may calculate the velocity on the fly in the regression formula:

> `log(q/acv) ~ ...`


## Time trend

Re-estimate the log-linear demand models for the two Tide products including a time trend. A time trend is a variable that proxies for the progress of time. Here, you can use the `week` variable as a time trend.

Explain why adding a time trend is important here. Discuss whether the demand estimates now make sense. Is there an improvement over the basic model specification?
```{r, fig.width = 6, fig.height = 4, fig.align = "center"}
tide128lm = glm(log(q_tide128/acv) ~ log(p_tide128) + log(p_tide64) +log(p_wisk64) + week, data = detergent_df)
summary(tide128lm)

tide64lm = glm(log(q_tide64/acv) ~  log(p_tide64) + log(p_tide128) + log(p_wisk64) + week, data = detergent_df)
summary(tide64lm)

detergent_df %>% 
  select("store":"p_tide64") %>% 
  pivot_longer(cols = q_tide128:p_tide64, names_to = c("variable", "product"), names_sep = "_") %>% 
  pivot_wider(id_cols = c(product, store, week, acv),
                   names_from = variable, values_from = value)  %>%
  mutate(velocity = q/acv) %>% 
  ggplot(data = ., aes(x = week, y = log(velocity), color = product)) + 
  geom_point(alpha = 0.5, size = 0.1) + 
  geom_smooth(method = "lm", se = FALSE)

```

Sales are not very clearly trending over time about tide128, but has clear impact about tide 64
The week coefficients have small enough P value to reject H0, but the estimate itself is very small.
Even with week 52, the log(velocity) is only 13%.

Now it makes more sense as a 1% increase in Wisk 64oz price will lead to 0.35% increase of Tide 64oz price (vs. a decrease of 0.87% in the previous model without time stamp). Now we can analyze elasticity based on an aligned timeframe, which helps us to eliminate misaligned information. It is an improvement compared to the basic model specification.

## Focus on non-promoted weeks

In the data, weeks where at least one product was promoted are flagged by the dummy variable `promoflag`. A value of 1 indicates a promoted week. In what fraction of store-weeks was at least one of the detergents promoted?

Now create (`filter`!) a new data set that only includes store-weeks in which none of the products were promoted.

Re-estimate the log-linear demand models with a time-trend for the two Tide products only using data from non-promoted store-weeks. Discuss whether the demand estimates now make sense. Is there an improvement over the previous model specification? Provide some intuition for the change in the estimated own-price effects.

```{r}
detergent_df_nonpromo = detergent_df %>% filter(promoflag == 0)

tide128lm = glm(log(q_tide128/acv) ~ log(p_tide128) + log(p_tide64) +log(p_wisk64) + week, data = detergent_df_nonpromo)
summary(tide128lm)

tide64lm = glm(log(q_tide64/acv) ~  log(p_tide64) + log(p_tide128) + log(p_wisk64) + week, data = detergent_df_nonpromo)
summary(tide64lm)

detergent_df %>% 
  select("store":"p_tide64") %>%
  filter(promoflag == 0) %>% 
  pivot_longer(cols = q_tide128:p_tide64, names_to = c("variable", "product"), names_sep = "_") %>% 
  pivot_wider(id_cols = c(product, store, week, acv),
                   names_from = variable, values_from = value)  %>%
  mutate(velocity = q/acv) %>% 
  ggplot(data = ., aes(x = week, y = log(velocity), color = product)) + 
  geom_point(alpha = 0.5, size = 0.1) + 
  geom_smooth(method = "lm", se = FALSE)

```

Fraction = 1-2676/14745=81.9%
Tide 128 analysis:
Notice Tide 64oz is not statistically significant in this case, which could be interpreted as changes in Tide 64oz price will not have a significant impact on Tide 128oz demand. This indicates limited cannibalization within Tide products.
Tide 64oz analysis:
Similar to the previous analysis, Tide 128oz is not statistically significant in this case, which could be interpreted as changes in Tide 128oz price will not have a significant impact on Tide 64oz demand. This indicates limited cannibalization within Tide products.
Interestingly, a 1% increase in Wisk 64oz price will lead to 0.52% decrease in Tide 64oz demand according to this model. Further investigation will need to be done on this issue. 
Intuitively, once the products are not on sale, customers tend to be less price sensitive on the base price. We need to analyze store fixed effects for a better model. Controlling promotion is not a substantial improvement of the model.

## Store fixed effects

Re-estimate the log-linear demand models for the two Tide products including a time trend and store fixed effects using the data for the non-promoted store-weeks. 




\medskip

Note: You can estimate models with fixed effects by explicitly adding dummy variables to the model. We have seen before (in the R introduction) that such dummy variables can be created in the regression model formula. In this example, the fixed effects are based on the store-identity, and hence the dummy variables can be added to the regression model formula in the form 

> `y ~ x1 + x2 + ... + factor(store)`

Although this approach works in our example I do not recommend it. The reason is that in other applications the number of dummy variables can be large, and then the approach of directly creating dummy variables will become computationally infeasible.

Instead, **install and use the lfe package**.

```{r}
library(lfe)
```

"lfe" stands for linear fixed effects. When using the lfe package you estimate a regression model using `felm` (fixed effects linear model) instead of `lm`. The use of the `felm` function is almost identical to the use of `lm`. The only difference is that you may add a variable that indicates the different groups for the fixed effects after the vertical bar `|`. In our example:

> `felm(log(q) ~ log(p1) + ... | store, data = ...)`

In general, you may use one or multiple fixed effects variables, such as

> `felm(y ~ x1 + x2 + ... | fe_var1 + fe_var2 + ..., data = ...)`

Note that when you display the regression output using `summary` the fixed effects are not shown. This is by design, because a large number of fixed effects makes the regression output virtually unreadable.

\medskip

When using store fixed effects you can simply use `log(q)` as dependent variable. Indeed, confirm that when using store fixed effects the elasticity and time trend estimates are identical irrespective of whether you use the log of the velocity or the log of units as dependent variable. Why is there not difference between the two specifications?

\medskip

Do the estimates reveal an improvement over the model specification before?

```{r}
tide128lm = felm(log(q_tide128/acv) ~ log(p_tide128) + log(p_tide64) +log(p_wisk64) + week | store, data = detergent_df_nonpromo)
summary(tide128lm)

tide64lm = felm(log(q_tide64/acv) ~  log(p_tide64) + log(p_tide128) + log(p_wisk64) + week | store, data = detergent_df_nonpromo)
summary(tide64lm)

#improve a lot about tide128 model but not much about tid64 model
```

The own price elasticity is smaller. The cross price elasticity of the Tide 128 oz to the Tide 64 oz is 0.9. The coefficient for the Wisk 64 oz is not statistically significant and we cannot derive any conclusions for the brand difference.
This looks like 128oz is cannibalizing 64oz Tide. But Tide 64oz price has no effect on Tide 128. Since now Tide 64oz has a relatively low elasticity we could consider increasing its price. 
There is no difference if we use log(q/acv) or log(q) because when we look at linear fixed effects, each store is its own category. Therefore the attempt to take into account of ACV in the modeling result is no longer necessary here.

\newpage




# Pricing and profitability analysis

Tide's retail margin at Dominick's is 25 percent, and P&G's marginal cost of producing Tide laundry detergent is 2.7 cents per oz.

Calculate **base (regular) prices**, using the data for the **non-promoted** store-weeks, as follows:
$$\text{base price of Tide 128 = mean of price of Tide 128 across non-promoted store-weeks}$$

Use a similar calculation for Tide 64.

Calculate the **base volume** as average yearly chain-level volume sales:
$$\text{base volume of Tide 128 = no. of stores} \times 52 \times \text{mean sales of Tide 128}$$

Use a similar calculation for Tide 64. Recall that there are 86 stores in the data set.

\medskip

1. Calculate the total new expected volume of Tide, i.e. the new volume of the 128 oz and 64 oz products, from the following price changes:

    (i) A simultaneous 5 percent increase in the prices of Tide 128 and Tide 64 
    (ii) A simultaneous 5 percent decrease in the prices of Tide 128 and Tide 64
    (iii) A simultaneous 5 percent increase in the price of Tide 128 and 5 percent decrease in the price of Tide 64 
    (iv) A simultaneous 5 percent decrease in the price of Tide 128 and 5 percent increase in the price of Tide 64

```{r}
margin = 0.25
vcoz = 0.027

bptide128 = mean(detergent_df_nonpromo$p_tide128)
bptide64 = mean(detergent_df_nonpromo$p_tide64)

bvtdie128 = length(unique(detergent_df$store))*52*mean(detergent_df$q_tide128)
print(bvtdie128)
bvtdie64 = length(unique(detergent_df$store))*52*mean(detergent_df$q_tide64)
print(bvtdie64)

tide128lm = felm(log(q_tide128) ~ log(p_tide128) + log(p_tide64) +log(p_wisk64) + week | store, data = detergent_df_nonpromo)
summary(tide128lm)
b128_128 = summary(tide128lm)$coef[1,1]
#since p_tide 64 p value is too large, reject it.


tide64lm = felm(log(q_tide64) ~ log(p_tide128) + log(p_tide64) +log(p_wisk64) + week | store, data = detergent_df_nonpromo)
summary(tide64lm)
b64_128 = summary(tide64lm)$coef[1,1]
b64_64 = summary(tide64lm)$coef[2,1]

#(i)
new_q_tide128 = bvtdie128*(1.05^b128_128)
print(new_q_tide128)
new_q_tide64 = bvtdie64*(1.05^b64_128 * 1.05^b64_64)
print(new_q_tide64)

#(ii)
new_q_tide128 = bvtdie128*(0.95^b128_128)
print(new_q_tide128)
new_q_tide64 = bvtdie64*(0.95^b64_128 * 0.95^b64_64)
print(new_q_tide64)

#(iii)
new_q_tide128 = bvtdie128*(1.05^b128_128)
print(new_q_tide128)
new_q_tide64 = bvtdie64*(1.05^b64_128 * 0.95^b64_64)
print(new_q_tide64)

#(iv)
new_q_tide128 = bvtdie128*(0.95^b128_128)
print(new_q_tide128)
new_q_tide64 = bvtdie64*(0.95^b64_128 * 1.05^b64_64)
print(new_q_tide64)

```
    (i) A simultaneous 5 percent increase in the prices of Tide 128 and Tide 64
    From the results in 3.4, the quantity change is shown.
    $$\%\Delta{Q_{Tide128}} = \varepsilon_{own} \times \%\Delta{P_{Tide128}} + \varepsilon_{cross_{Tide64}} \times \%\Delta{P_{Tide64}} + \varepsilon_{cross_{Wisk64}} \times \%\Delta{P_{Wisk64}} \\ + \beta_{Week} \times \Delta{Week} + \text{Store Effect}$$
    $$\%\Delta{Q_{Tide64}} = \varepsilon_{own} \times \%\Delta{P_{Tide64}} + \varepsilon_{cross_{Tide128}} \times \%\Delta{P_{Tide128}} + \varepsilon_{cross_{Wisk64}} \times \%\Delta{P_{Wisk64}} \\ + \beta_{Week} \times \Delta{Week} + \text{Store Effect}$$
    With all others fixed, a simultaneous 5% price increase for Tide 128 and Tide 64 would generate a sales change of,
    $$\%\Delta{Q_{Tide128}} = -2.3836 \times 5 + 0.2097 \times 5 = -10.87 (\%)$$
    $$ \text{Volume} = BaseVolume \times (1 + \%\Delta_{Q_{Tide128}}) = 363,197 \times (1-10.87\%) = 323,719$$
    $$\%\Delta{Q_{Tide64}} = -1.4867 \times 5 + 0.9028 \times 5 = -2.92 (\%)$$
    $$ \text{Volume} = BaseVolume \times (1 + \%\Delta_{Q_{Tide64}}) = 327,717 \times (1-2.92\%) = 318,149$$
    \medskip
    (ii) A simultaneous 5 percent decrease in the prices of Tide 128 and Tide 64
    With all others fixed, a simultaneous 5% price decrease for Tide 128 and Tide 64 would generate a sales change of,
    $$\%\Delta{Q_{Tide128}} = -2.3836 \times (-5) + 0.2097 \times (-5) = 10.87 (\%)$$
    $$ \text{Volume} = BaseVolume \times (1 + \%\Delta_{Q_{Tide128}}) = 363,197 \times (1+10.87\%) = 402,674$$
    $$\%\Delta{Q_{Tide64}} = -1.4867 \times (-5) + 0.9028 \times (-5) = 2.92 (\%)$$
    $$ \text{Volume} = BaseVolume \times (1 + \%\Delta_{Q_{Tide64}}) = 327,717 \times (1+2.92\%) = 337,284$$
    \medskip
    (iii) A simultaneous 5 percent increase in the price of Tide 128 and 5 percent decrease in the price of Tide 64
    With all others fixed, a simultaneous 5% price increase for Tide 128 and 5% price decrease for Tide 64 would generate a sales change of,
    $$\%\Delta{Q_{Tide128}} = -2.3836 \times 5 + 0.2097 \times (-5) = -12.97 (\%)$$
    $$ \text{Volume} = BaseVolume \times (1 + \%\Delta_{Q_{Tide128}}) = 363,197 \times (1-12.97\%) = 316,103$$
    $$\%\Delta{Q_{Tide64}} = -1.4867 \times 5 + 0.9028 \times (-5) = -11.95 (\%)$$
    $$ \text{Volume} = BaseVolume \times (1 + \%\Delta_{Q_{Tide64}}) = 327,717 \times (1-11.95\%) = 288,563$$
    \medskip
    (iv) A simultaneous 5 percent decrease in the price of Tide 128 and 5 percent increase in the price of Tide 64
    With all others fixed, a simultaneous 5% price decrease for Tide 128 and 5% price increase for Tide 64 would generate a sales change of,
    $$\%\Delta{Q_{Tide128}} = -2.3836 \times (-5) + 0.2097 \times 5 = 12.97 (\%)$$
    $$ \text{Volume} = BaseVolume \times (1 + \%\Delta_{Q_{Tide128}}) = 363,197 \times (1+12.97\%) = 410,291$$
    $$\%\Delta{Q_{Tide64}} = -1.4867 \times (-5) + 0.9028 \times (-5) = 11.95 (\%)$$
    $$ \text{Volume} = BaseVolume \times (1 + \%\Delta_{Q_{Tide64}}) = 327,717 \times (1+11.95\%) = 366,870$$


2. Calculate the total new expected profits for each of the price changes in 1. Are the prices of Tide approximately optimal, or do you recommend changes to the product-line pricing of Tide?

```{r}
profit_tide128= 0
profit_tide64 = 0
total_profit = profit_tide128 + profit_tide64

#base
profit_tide128 = bvtdie128 * (bptide128*(1 - margin) - vcoz*128)
profit_tide64 = bvtdie64 * (bptide64*(1 - margin) - vcoz*64)
profit_tide128
profit_tide64
total_profit = profit_tide128 + profit_tide64
total_profit
#Base profit 1568306


pc_tide128 = -0.05
pc_tide64  = 0.05
#new_q_tide128 = bvtdie128*((1 + pc_tide128)^b128_128)
#new_q_tide64 = bvtdie64*((1 + pc_tide128)^b64_128 * (1 + pc_tide64)^b64_64)

profit_tide128 = bvtdie128*((1 + pc_tide128)^b128_128) * (bptide128*((1 + pc_tide128) *(1 - margin)) - vcoz*128)

profit_tide64 = bvtdie64*((1 + pc_tide128)^b64_128 * (1 + pc_tide64)^b64_64) * (bptide64*((1 + pc_tide64) *(1 - margin)) - vcoz*64)

total_profit = profit_tide128 + profit_tide64
total_profit

#(i)
#total profit is 1593502

#(ii)
#total profit is 1534799

#(iii)
#total profit is 1560281

#(iv)
#total profit is 1565150

```

\newpage




# Summarize your findings

(1) What is the extent of cannibalization within the Tide product line? 

Tide 128 price change impact Tide 64 salse, however, Tide 64 price wouldnot impact Tide 128 sales

(2) Does Tide face a competitive threat from Wisk?

Yes, Wisk 64 competete with Tide 128, however, the model is not enough to analysis competition about Tide 64 since the P value is too large (>10%)

(3) How do you evaluate the current pricing tactics? Do you recommend changes?

Current pricing is not optimal for short term profit. If company doesn't focus on a market penetration strategy, I suggest rise up price for both product.
Based on our observation in Q.4, increasing Tide 64oz and Tide 128oz price by 5% offers the highest profit; this indicates that in order to maximize profit, we can consider increase prices.

